Highly Available Transactions: Virtues and Limitations

### 2013-03-20 writeup/futurework.tex

\section{Future Work}

VII. Future Work (.75pg)

* What properties are available?
	* We've focused on ACID Isolation, but there are lots of other models to consider?
	* Which consistency criteria? Which models? There are infinite, so how do we choose?
	* Set ourselves up for APEs

* Full exploration of HAT space
	* Many parameters; $2^10$ models at the least
	* Need to mitigate the cost of tracking causality with high availability
		* The cost moving from S-HA (like Eiger) to truly HA (R-HA) seems to be high
		* e.g., Eiger relies on metadata GC to keep writes small; is O(|keys|) in limit
			* Not meant as a fault--it's optimized for a particular scenario--but curious to consider

* Revisit semantics-based concurrency control
	* Could taxonomize properties like escrow
	* Can also consider models with SLAs and bounded messaging asynchrony
	* Our focus is on "classic" distributed systems trade-off
		* However, lots of interesting hybrids (maybe cite Doug Terry's recent stuff)

>>>>

\section{Future Work}
\label{sec:futurework}

\textbf{What is the space of consistency and isolation properties?} In
this work, we have taxonomized several well-defined and
well-documented consistency properties. This is useful as it
contextualizes decades of existing database design and algorithmic
techniques but does not provide guidance as to what consistency models
are necessarily the ``strongest''---or admit the fewest number of
executions---within each availability class. While a recent report
shows that no model stronger than causal consistency can be achieved
in a sticky highly available, one-way convergent system~\cite{cac},
there are infinite highly available and sticky highly available models
to consider. For example, always returning the value $1$ in response
to reads to any data item is incomparable to (neither stronger nor
weaker than) causal consistency: returning $1$ does not respect
causality, and causal consistency returns values other than
$1$. Always returning the value $2$ is similarly incomparable, and, as
an example, we can enumerate models that return all integers and real
numbers--an uncountably infinite set--along with models that respond
to every tenth reply and so on.

The current state of the art places several useful models within
various categories in the availability space. This is still a
practical result---take, for instance, CAP's impact on distributed
systems of the last decade. But what property defines the boundaries
between classes of highly available systems? Is there a meta-property
that exactly characterizes what properties are achievable with high
availability or sticky high availability? This is an area of ongoing
research within the HAT project.

\textbf{What are the costs and benefits within HAT designs?} There are
many possible combinations of HAT guarantees, and the performance and
ease of implementation of a HAT system will likely vary depending on
which guarantees it attempts to provide. As we have alluded to, the
cost in moving from a sticky high available system to a truly highly
available system seems expensive. As a concrete example, the Eiger and
COPS systems use metadata garbage collection and assume linearizable
clients within each datacenter in order to keep the overhead of
providing causal consistency small. Without performing extensive
client-side caching, these techniques do not seem applicable to a
highly available implementation. This is not intended as a fault of
the prior systems---they perform well for the task at hand---but hint
at the complexity of moving between sticky and truly highly available
systems.

\textbf{Where do semantics-based concurrency control lie in the HAT
  landscape?} The vast amount of literature on semantics-based
concurrency control provides hints as to when coordination can be
avoided but still providing ``strongly consistent'' outcomes. We have
performed a narrow case study in this work, but a general
taxonomization of techniques such as escrow transactions and Sagas
could help answer the challenge of programming in a weakly consistent
environment. These techniques might be aided by considering models
with service-level agreements or bounded messaging asynchrony. While
our focus is on ``classic'' distributed systems trade-offs, there are
several interesting hybrids to consider.

### 2013-03-20 writeup/hat-analysis.tex

We can consider two variants of Cut Isolation. The first is relatively
straightforward, and reasons about cuts over discrete data items. We
will call this \textit{Item Cut Isolation}. The second is known as
preventing the \textit{Phantom Problem}, whereby two successive
predicate-based reads return different data. We call predicate-based
cut isolation, which prevents Phansoms, \textbf{Phantom Cut
  Isolation}.

@@

Read-your-writes can be accomplished via client-side caching and
alternatively comes ``for free'' when requests are routed to a sticky
set of servers. Achieving causal consistency requires maintaining
read-your-writes along with the prior session guarantees. Both are
reasonable strategies often employed in real-world
systems~\cite{vogels-defs}, albeit at the cost of true high
availability or client-side storage.

@@

\textbf{}

* Recency guarantees on data items
	* CAP

@@

@@

\begin{table}
\begin{tabular}{| c | p{6cm} | }\hline
HA & Transactional Atomicity (TA), Read Uncommitted (RU), Read
Committed (RC), Item Cut Isolation (P-CI), Phantom Cut Isolation
(P-CI), Monotonic Reads (MR), Monotonic Writes (MW), Writes Follow
Reads (WFR)\\\hline Sticky-HA & Read Your Writes (RYW), Causal
Consistency\\\hline Unavailable & Cursor Stability (CS)\lostupdate,
Read Consistency (CR)\lostupdate, Snapshot Isolation (SI)\lostupdate,
Repeatable Read (RR)\lostupdate\rwskew, Serializability
(S)\lostupdate\rwskew, Causal
Serializability\lostupdate\rwskew, Strict
Serializability\lostupdate\rwskew\linearizable \\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing write skew\rwskew, and requiring
  linearizability\linearizable.}
\label{table:hatcompared}
\end{table}

@@

 \node[draw=none,fill=none] (ici) at (1.2, 1.6) {\underline{I-CI}};
 \node[draw=none,fill=none] (pci) at (1.2, 2.4) {\underline{P-CI}};
 \node[draw=none,fill=none] (rc) at (1.2, .8) {\underline{RC}};
 \node[draw=none,fill=none] (ru) at (1.2, 0) {\underline{RU}};

 \node[draw=none,fill=none] (ta) at (0, 0) {\underline{TA}};
 \node[draw=none,fill=none] (mr) at (2.4, 0) {\underline{MR}};
 \node[draw=none,fill=none] (mw) at (3.6, 0) {\underline{MW}};
 \node[draw=none,fill=none] (wfr) at (4.9,0) {\underline{WFR}};
 \node at (6.3,0) [sticky] (ryw) {\underline{RYW}};

 \node at (3.6, 2) [sticky] (causal) {\underline{Causal}};
 \node[noha] (cs) at (2, 2.8) {CS};
 \node[noha] (rr) at (.6, 3.0) {RR};
 \node[noha] (si) at (-.4, 2.4) {SI};
 \node[noha] (cr) at (-.4, 1.2) {CR};
 \node[noha] (s) at (-.4, 3.5) {S};
 \node[noha] (causal-s) at (3.6, 3.5) {Causal S};
 \node[noha] (strict-s) at (6, 3.5) {Strict S};

@@

 \draw [->, blue] (mr) -- (causal);
 \draw [->, blue] (mw) -- (causal);
 \draw [->, blue] (wfr) -- (causal);
 \draw [->, blue] (ryw) -- (causal);

@@

 \draw [->, red] (ici) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (pci) -- (si);
 \draw [->, red] (rr) -- (s);
 \draw [->, red] (si) -- (s);
 \draw [->, red] (ta) -- (cr);
 \draw [->, red] (rc) -- (cr);
 \draw [->, red] (cr) -- (si);
 \draw [->, red] (ta) -- (rr);
 \draw [->, red] (s) -- (causal-s);
 \draw [->, red] (causal) -- (causal-s);
 \draw [->, red] (causal-s) -- (strict-s);

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model strength,
  while all underlined models can be combined (e.g., the powerset of
  the set of underlined models are all valid models). The combination
  of all underlined, non-boxed models represents maximum ACID HAT
  semantics, while the combination of underlined and boxed models
  represents maximum ACID sticky HAT models.}
\label{fig:hatcompared}
\end{figure}

@@

TRASH

We can accomplish these three requirements in two ways. First, clients
can locally cache their reads and writes, updating caches when they
read admissible replacements from
replicas~\cite{sessionguarantees}. Second, clients can ensure
\textit{stickiness} with groups~\cite{vogels-defs} (subject to the
same groups caveat as before).  For monotonic writes, clients should
generate increasing IDs for each transaction they perform. Caching and
ID generation are both available.

To provide writes follow reads,
severs (or clients) can buffer transaction writes until the
transactions the writes depend on have been written to all
replicas~\cite{cops}. If we consider a group model, each group can
independently apply transactions in order via a (possibly sharded)
log-shipping approach~\cite{causalmemory, cops, eiger, swift}.

What about existing systems?
* Database technology developed for single-node systems; gold standard: serializability
	* Serializability is not actually highly available; give an example!
	* Much of the database literature presumes serializability and does not consider high availability.
	* For the literature that doesn't presume serializability, it's not presented in a HA context?
* Consequence: traditional systems are not optimized for high availability
	* Consider two-phase locking in a distributed environment
	* As we will see (forward reference to Evaluation), many existing transactional systems encounter similar difficulties
* Key question in this paper: What transactional semantics are highly available, and which aren't?
	* Side note: there are infinite incomparable consistency models (e.g., always return 2, always return 3, …)
	* Our goal is to unify distributed systems literature with ACID transaction model.
* Argument: Highly Available Transactional Systems (HATS) require rethinking existing models. This work is a first step.

In this sense, HATs are similar to RAID: optimizing for graceful
handling of a worst-case failure scenario improves average-case
performance.

IV. HATS and ACID (4pg)

We will show what semantics are available in HATS as well as which aren't. Our strategy will be to build up a set of properties until we reach a point of unavailability. We will start with isolation and atomicity since they are well-defined in the literature and most meaty.

A. *What's Available in Isolation and Atomicity?*

* We can prevent Dirty Write phenomena by consistently ordering transactions
	* This is basically a cross-data item convergency property, as are found in eventually consistent systems!
* Read Committed
	* Don't read dirty data by never exposing dirty data to readers!
* ANSI Repeatable Read => rename to cut isolation
	* Repeatable Read is a tricky guarantee. In ANSI SQL Spec, rather weak, but in follow-on work, it's pretty strong.
	* The property that the ANSI SQL spec talks about is about reading a "snapshot" of data items, along with your own writes.
		* A "snapshot" describes a "consistent cut" across data items; we'll call it "cut isolation"--when paired with below guarantees, makes sense.
	* This doesn't provide any constraints on writes.

* There are several properties that aren't discussed in the formal literature on isolation guarantees but which are really useful.

* Transactional atomicity: "A" in ACID
	* Once a transaction reads one of another transactions's writes, its subsequent reads will return the transactions's other writes (or a suitable "later" write).
	* Be really pedagogical here: "transactional atomicity" versus "distributed linearizability" both called "atomicity"; we actually explore the differences later, in Discussion section.
	* Talk about implementation here: it's basically uniform eager reliable delivery from distributed systems literature!
	* Note that this doesn't make recency guarantees (forward reference to later section on "impossibility")

* Session guarantees give useful guarantees within and across transactions
	* Without them, cut isolation is sort of meaningless: we can always just return NULL!
	* Several properties: monotonic writes, monotonic reads, writes follow reads, read your writes
	* We can define them within a transaction only (e.g., transactional monotonic writes) or across subsequent transactions by the same client (e.g., client monotonic writes).

* Session guarantees are tricky
	* You can implement all but read-your-writes in a R-HA system
		* Never show a write until all replicas in the system have seen it!
		* This means that, if a client switches replicas, then the replica will have a satisfying set of writes.
                * Take S3 as an example of a system where stickiness coudl have helped, but no one really talked about it!
		* Downside: when do writes become visible? Only when all replicas have seen it!
		* Side note: assuming that, if replicas can enter and leave, they can only become active once they "catch up" to global lower bound
			* We leave dynamic replica selection as Future Work (N.B. in APEs!)
	* Read-your-writes requires S-HA
		* Example: partition a client from all but one replica, have the client issue a write, which has to commit. Next, lift the partition between the client and the other replicas and partition the client and its previous replica. It can't read it's own write!
		* This also means that causality, which is MR+MW+WFR+RYW is unavailable!
		* S-HA is *doable* but it does result in substantially lowered unavailability unless one caches (e.g., bolt-on causal consistency)

\noindent\textit{Groups:} We can optimize write visibility delay by
partitioning servers and clients into groups, such that each group
contains a fully replicated set of data items and all clients within a
group contact only the servers in the group. Once all transaction
values are present within a group, values can be installed as
$d_{good}$ (as opposed to once they are present on all replicas). For
example, in a multi-datacenter setting, all clients and servers within
a datacenter may form a group. However, group-based atomicity is not
highly available according to our definition unless clients and
servers in a group fail together, as is often assumed in a
geo-replicated context~\cite{cops, eiger}.

* What do we have?
	* In Adya, we get up to Causal+Cut Isolation for S-HA, and PL-2 for R-HA.
	* In distributed systems terminology, in S-HA we have causal consistency but each transactions' updates are a cycle in the happens-before graph!
		* Stress that this is a big unification of the existing models.
	* Recap, pedagogically, why this is highly available and what this means: no locks, no central coordination, no need for RTTs; bring back to Section II.

B. *What's not available in Isolation?*

Now, at least in terms of the existing literature, we hit the limits.

Let's sketch out a few conditions, then we'll discuss them later:

Now refer to Adya's chart. No PL-SI, PL-SS, PL-2.99, etc.

C. *What's available in consistency?*

* In general, we can't maintain correctness conditions over arbitrary data items.
	* This is due to serializability problems.
	* Uniqueness, for example, is out the window

* We can evaluate locally checkable correctness criteria (e.g., check for null)
	(QUESTION: should we save this for APEs?)

* However, with semantic knowledge, we can do a little bit better
	* Commutative updates and logical monotonicity are fine; e.g., write-write conflicts don't matter

Example: In TPC-C New Order, the new order id assignment isn't commutative, tough. but the inventory checking *is*!

D. Durability

* If you want data to survive *F* failures, you'll have to replicate to *F* replicas.
	* This is pretty easy.

E. Summary

* Left with a bunch of binary properties: 
R-HA: Prevent Dirty Write, Read Committed, Cut Isolation, Transactional Atomicity, Transactional Monotonic Reads, Transactional Monotonic Writes, Transactional Writes-Follow-Reads, Client Monotonic Reads, Client Monotonic Writes, Client Writes-Follow-Reads
S-HA: Client Read Your Writes, Transactional Read Your Writes

* Combining all of the above results in the strongest models we've yet seen. No

*Discussion*

Stickiness and PNUTS
* As an example of a real system going from "CP" to "AP", [PNUTS dropped per-record mastering](\url{http://developer.yahoo.com/blogs/ydn/posts/2010/06/sherpa_update/#4})

* Composition of Properties can be tricky:
	* Consider ensuring RR and TA:
		* If you do this naively, you may end up returning lots of nulls
		* Give example: T1 w(x1) w(y1) T2 w(x2) w(y2) T3 r(x) r(y) => T3 better read x1, y1 or x2, y2!
	* Some are more expensive; not likely to achieve causality with less than O(clients) metadata required
		* Note that making stronger assumption about clusters, like linearizability *can* help (but then not HA!)

* Transactional Atomicity versus Linearizability
	* What's the difference between our TA and distributed systems linearizability? Is there some sort of mapping?
		* TA: writes to multiple keys are indivisible to readers
		* Linearizability: writes to single key on multiple servers are indivisible to readers
	* The differences are two-fold:
		* In linearizability (and safe and regular registers), writes are visible to clients immediately after they finish
			* HATS have no recency requirement
		* In linearizability, all clients see all writes at the same time!
			* In TA, clients see writes at different times depending on what replicas they contact
			* No such analog in traditional distributed systems literature on replica consistency

* S-HA Monotonic Reads and Monotonic Writes
	* In the model we discussed, you can get monotonic reads in a R-HA system by waiting until all replicas see a write.
	* This leads to poor visibility--writes take a long time to show up. Even if you didn't want to read your writes, you might want to read other peoples' writes!
	* Accordingly, there are valid reasons for S-HA session guarantees, namely, visibility.

>>>>

We can consider two variants of Cut Isolation. The first is relatively
straightforward, and reasons about cuts over discrete data items. We
will call this \textit{Item Cut Isolation}. The second is known as
preventing the \textit{Phantom Problem}, whereby two successive
predicate-based reads return different data. We call predicate-based
cut isolation, which prevents Phansoms, \textbf{Phantom Cut
  Isolation}.\footnote{Oracle provides an isolation level called
  ``Statement Level Read Consistency''; this is analogous to Phantom
  Cut Isolation at the level of a single statement. We do not consider
  this isolation model here as it is subsumed by standard Snapshot
  Isolation but we believe it is easily adaptable.}

@@

Read-your-writes can be accomplished via client-side caching and
alternatively comes ``for free'' when requests are routed to a sticky
set of servers. Achieving causal consistency requires maintaining
read-your-writes along with the prior session guarantees. Both are
reasonable strategies often employed in real-world
systems~\cite{vogels-defs}, albeit at the cost of true high
availability or client-side storage. Perhaps interestingly,

@@

The inability to prevent Lost Update means that Consistent Read,
Snapshot Isolation, and Cursor Stability are all unavailable. The
inability to prevent Lost Update or Write Skew means that Repeatable
Read, Conflict Serializability, and One-Copy Serializability are
unavailable.

\subsubsection{Unavailable Recency Guarantees}

Data storage engines make various recency guarantees on reads of data
items. One of the most famous is linearizability, which states that
reads will return the last completed write to a data item, and there
are several other (weaker) variants such as safe and regular register
semantics. When applied to transactional semantics, the combination of
one-copy serializability and linearizability is called \textit{strong
  (or strict) one-copy serializability} or \textit{externally
  consistent transactions} (e.g., Spanner). It is also common,
particularly in systems that allow reading from masters and slaves, to
provide a guarantee such as ``read a version that is no more than five
seconds out of date'' or similar. Unfortunately, as the CAP Theorem
demonstrates, linearizability is unavailabile, and any indefinitely
long partition can cause violation of any recency bound.

@@

We summarize our results in Table~\ref{table:hatcompared}. A wide
range of isolation levels are achievable in a highly available system,
including transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve PRAM
consistency, read your writes guarantees, and causal
consistency. However, many other prominent models, such as Snapshot
Isolation, One-Copy Serializability, and Strict Serializability cannot
be achieved due to the inability to prevent Lost Update and Write Skew phenomena.

We illustrate the hierarchy of available, sticky available, and
unavailble consistency models we have discussed in
Table~\ref{fig:hatcompared}. Many models are simultaneously
achievable: for example, a HAT system can provide monotonic reads and
transactional atomicity or transactional atomicity and read your
writes and, for maximum semantic richness, can provide all of
Transactional Atomicity, Predicate Cut Isolation, Monotonic Reads,
Monotonic Writes, and Writes Follow Reads. An unavailable system might
provide PRAM and Serializability, as advocated by Salem and
Daudjee~\cite{needed}.

To the best of our knowledge, this is the first unification of
database ACID, distributed consistency, and session guarantee
models. Perhaps satisfactorily, strong one-copy serializability
subsumes all other models, while considering the (large) power set of
all compatible models (e.g., only considering client-centric session
consistency guarantees, the diagram depicts 32 possible HAT
combinations) hints at the vast expanse of consistency models found in
the literature. This taxonimization is \textit{not} exhaustive, but we
believe it lends substantial clarity into the relationships between a
large subset of the prominent ACID and distributed consistency
moels. Additional models of read/write transactions that we have
omitted should be easily classifiable based on the available
primitives and unavailable anomalies we have already discussed.

@@

\begin{table}
\begin{tabular}{| c | p{6cm} | }\hline
HA & Transactional Atomicity (TA), Read Uncommitted (RU), Read
Committed (RC), Item Cut Isolation (P-CI), Predicate Cut Isolation
(P-CI), Monotonic Reads (MR), Monotonic Writes (MW), Writes Follow
Reads (WFR)\\\hline Sticky-HA & Read Your Writes (RYW), PRAM,
Causal\\\hline Unavailable & Cursor Stability (CS)\lostupdate,
Snapshot Isolation (SI)\lostupdate, Repeatable Read
(RR)\lostupdate\rwskew, Conflict Serializability
(CSR)\lostupdate\rwskew, One-Copy Serializability
(1SR)\lostupdate\rwskew, Recency\linearizable, Safe\linearizable,
Regular\linearizable, Strong 1SR\lostupdate\rwskew\linearizable
\\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing write skew\rwskew, and requiring
  recency guarantees\linearizable.}
\label{table:hatcompared}
\end{table}

@@

 \node[draw=none,fill=none] (ici) at (1.2, 1.6) {I-CI};
 \node[draw=none,fill=none] (pci) at (1.2, 2.4) {P-CI};
 \node[draw=none,fill=none] (rc) at (1.2, .8) {RC};
 \node[draw=none,fill=none] (ru) at (1.2, 0) {RU};

 \node[draw=none,fill=none] (ta) at (0, 0) {TA};
 \node[draw=none,fill=none] (mr) at (2.4, 0) {MR};
 \node[draw=none,fill=none] (mw) at (3.6, 0) {MW};
 \node[draw=none,fill=none] (wfr) at (6.3,0) {WFR};
 \node at (5,0) [sticky] (ryw) {RYW};

 \node[noha](recency) at (7.7, 0) {recency};
 \node[noha](safe) at (7.7, 1) {safe};
 \node[noha](regular) at (7.7, 2) {regular};
 \node[noha](linearizable) at (7.7, 3) {linearizable};
 \node at (3.6, 2) [sticky] (causal) {causal};
 \node at (3.6, 1) [sticky] (pram) {PRAM};
 \node[noha] (cs) at (2, 2.8) {CS};
 \node[noha] (rr) at (.6, 3.0) {RR};
 \node[noha] (si) at (-.4, 2.4) {SI};
 \node[noha] (s) at (-.4, 3.5) {CSR};
 \node[noha] (sr) at (2.2, 3.6) {1SR};
 \node[noha] (ssr) at (4.95, 3.7) {Strong-1SR};

 \draw [->, red] (recency) -- (safe);
 \draw [->, red] (safe) -- (regular);
 \draw [->, red] (regular) -- (linearizable);
 \draw [->, red] (linearizable) -- (ssr);
 \draw [->, red] (sr) -- (ssr);

@@

 \draw [->, blue] (mr) -- (pram);
 \draw [->, blue] (mw) -- (pram);
 \draw [->, blue] (wfr) -- (causal);
 \draw [->, blue] (ryw) -- (pram);
 \draw [->, blue] (pram) -- (causal);

@@

 \draw [->, red] (ici) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (pci) -- (si);
 \draw [->, red] (rr) -- (s);
 \draw [->, red] (si) -- (s);
 \draw [->, red] (ta) -- (si);
 \draw [->, red] (ta) -- (rr);
 \draw [->, red] (s) -- (sr);
 \draw [->, red] (causal) -- (sr);

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model
  strength. Models that do not share a common ancestor can be
  simultaneously achieved, and the resulting availability is that of
  the weakest available model in the combination.}
\label{fig:hatcompared}
\end{figure}

@@

In this section, we discuss several subleties we have mostly elided up
to this point, specifically addressing model composition,
transactional atomicity versus linearizability, and stickiness
requirements.

\noindent\textbf{Model Composition} may be non-trivial. Consider the following transactions:
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(1)
\\T_3 &: r_x(a)~r_y(b)
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and these are the only transactions that are executed by the system,
then $T_3$ needs to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means
that either the implementation should frequently return null
(definitely undesirable and possibly non-convergent), keep multiple
versions of each data item (necessitating potentially complicated
distributed garbage collection), or use pre-declared read sets to
fetch a consistent cut of keys before the transaction begins to
execute. Using client-side caching can alleviate some of these
challenges, but then the system becomes sticky high available.

Composition cost varies by property. For instance, Charron-Bost has
proven that, to capture causality between $N$ communicating processes,
standard vector-based approaches face an upper bound of $O(N)$. This
means that, with $100,000$ clients, each write might be accompanied by
$100,000$ vector timestamps. This is difficult to scale. By
compromising on availability (e.g., treating a datacenter as a
linearizable cluster), this overhead can be reduced, but it is much
cheaper to provide, say, read your writes, compared to causality.

\noindent The relationship between \textbf{Transactional Atomicity and
  Linearizability} is not immediately clear. One might attempt to find
a mapping between the two: Transactional Atomicity dictates that
writes to multiple keys across multiple servers are made visible to
readers all at once, while Linearizability dictates taht writes to a
single key on multiple servers are made visible to all readers at
once. The difference is two-old. First, in linearizable (and safe and
regular) systems, writes are made visible to all clients immediately
after they finish. With Transactional Atomicity, there is no recency
requirement. Second, in linearizabile systems, all clients see all
writes at the same time. With Transactional Atommicity, clients see
writes at different times depending on which relicas they contact. We
are not aware of a comparable analog in the distributed systems
literature. Accordingly, despite apparent similarities, Transactional
Atomicity is much weaker than Linearizability.

\noindent There is a substantial \textbf{visibility benefit to sticky
  availability}: writes can safely become visible to readers much more
quickly in a sticky available system. In the model we discussed, it is
possible to achieve several properties like monotonic reads in a
highly available system by waiting until all servers have seen a
write. However, this leads to horrible visibility---new writes will
not become visible to clients in the presence of partitions. A client
that does not want to guarantee read-your-writes (due to the sticky
availability requirement) may still wish to read other clients' writes
with timeliness. Accordingly, a system may forgo high availability in
order to provide better visibility. Alternatively, a system may
enforce stickiness in the absence of partitions, then ``lose'' session
guarantees by having clients reconnect to an available replica. This
latter strategy provides session guarantees between partitions, and
each partition event may cause a ``regression'' in the desired order
of partitions, albeit at the benefit of availablilty.

It is increasingly common in the practically-oriented distributed
systems literature to assume that clients will be able to remain
sticky with servers within its datacenter. More generally, if
client-server sticky pairs are partitioned into disjoint sets, or
groups, then additional guarantees like transactional atomicity can be
accomplished with better visibility: writes can be made visible once
they are present on all other servers within a group (as opposed to
all servers in the system).

### 2013-03-20 writeup/intro.tex
More writing

Many recent systems have eschewed ACID guarantees.

CAP Theorem says you can't have consistency and availability and also talks about partitions.

What does CAP have to say about ACID? Recent workshop result on Highly Available Transactions briefly demonstrated that some ACID guarantees are indeed achievable in a highly available manner.

Many ACID systems were designed either i.) for single systems or ii.)
to provide serializability. The combination of weak isolation and
distributed systems necessitates further study and
design. \textit{What is the benefit of high availability and what is
  the cost?}

Highly Available Transactions are the class of multi-object operations
and their respective semantics that can be provided with high
availability.

Weak isolation has a cost: in the words of one database pioneer, you
``fall off a cliff'' when it comes to reasoning about your
program. However, it also has a benefit: availability and latency.

Our goal in this paper is to objectively map the space of trade-offs and

 This paper fills in the remaining gaps.

What does high availability even mean in a transaction environment?

What models are achievable, and which aren't?

@@

\begin{itemize}
\item We present a model for high availability in a transactional environment, including Replica-agnostic High Availability (R-HA) and Sticky High Availability (S-HA).
\item We analyze the existing literature on ACID properties (particularly, ANSI SQL Spec, Adya, Gray) to show which properties* with a focus on isolation* are achievable in HATS. Achievable properties include transactional atomicity, variants of repeatable read isolation, and session guarantees. We also describe unavailable properties, like preventing Lost Update and Read and Write Skew. We believe that this is the first such characterization in the literature.
\item We further motivate the study of HATS via the analysis of existing concurrency control algorithms in a highly available environment as well as through deployment of an experimental database prototype on public cloud infrastructure.
\end{itemize}

>>>>

Requirements for high availability, predictable low latency, and
simplicity of coordination protocols have led many recent distributed
database designs to renounce many guarantees provided by traditional
database systems. In particular, many systems forgo providing ACID
guarantees, which often provide single-system image (SSI)
programmability---at the provable expense of increased coordination
and, by proxy, unavailability. The formal introduction of the CAP
Theorem, which proves the incompatiblity of ``always on'' operation
with guarantees on data item recency in the presence of unreliable
networks has led to a large number of systems prioritizing high
availability, which brings with it guaranteed low latency, which is especially
important in geo-replicated contexts. Accordingly, we observe a dearth
of systems providing general-purpose multi-operation
\textit{transactions} over multiple data items with useful semantics.

The relationship of the CAP Theorem with ACID semantics is poorly
understood. While serializability---the gold standard of SSI
transactions---is unavailable, databases have a long tradition of
providing weaker isolation and consistency guarantees in order to
provide greater concurrency and higher performance. While these weaker
guarantees do not provide SSI for arbitrary general-purpose read/write
transactions, they are widely employed (Section~\ref{needed}),
overwhelmingly as the default setting in today's traditional ``ACID''
and up-and-coming ``NewSQL'' databases and often as the maximum
guarantees offered. While the prevalence of these weak isolation
configurations is encouraging, the current literature does not provide
guidance as to the operation of weakly consistent isolation guarantees
in a distributed context. Rather, we have a strong understanding of
weak isolation in the single-server context from which they originated
as well as several techniques for providing distributed
serializability but little to no guidance as to which ACID guarantees
are compatible with Highly Available Transactions (HATS), or can be
provided with high availability and low latency.

This work attempts to bridge this gap between ACID properties and
highly available systems. We analyze a range of existing ACID and
distributed consistency models in order to provide a formal
taxonomization of which models are achievalbe in a HAT system.  While
several guarantees, like serializability and Snapshot Isolation, are
unachievable due to reliance on preventing Lost Update and Write Skew
phenomena, many other ACID guarantees such as providing Read Committed
isolation, preventing Fuzzy Read phenomena, and ensuring atomicity of
transactional updates are achievable. We also attempt to bridge the
often distant worlds of distributed system consistency and ACID
database consistency by analyzing the relationship between so-called
session guarantees and ``register'' semantics. In performing this
taxonomization, we also introduce a hierarchy of highly available
system guarantees as, for example, some guarantees require
``stickiness,'' or fate-sharing between clients and database
servers. We believe that this analysis is the first that has been
performed over such a broad range of ACID and distributed consistency
models.

While this model classification is useful, we seek to understand the
implications of choosing to employ high availability. HATs provide
high availability and low latency but suffer from weakened semantic
guarantees. Traditional ACID properties provide strong semantic
guarantees but do not provide high availability or low latency. We
attempt to objectively study these virtues and limitations by
performing an extensive survey of practitioner accounts and research
literature, performing experimental analyses on modern cloud
infrastructure, and analyzing representative applications for their
semantic requirements. While our results are not definitive, they
suggest that HATs offer a one to three order of magnitude latency
decrease compared to traditional protocols, while providing acceptable
semantics for programs with monotonic and commutative
updates. Similarly, HAT guarantees can be used to ensure that
arbitrary foreign key constraints are satisfied but not necessarily
uniqueness constraints. For programs with non-monotonic transactional
requirements, HATs may fall short, necessitating greater coordination
strategies. Speed and availability are not without a price, but we
attempt to estimate the size of the tag. Succintly, \textit{what is
  benefits do highly available transactions offer and what is their
  cost?}

@@

\begin{itemize}
\item We present a model for high availability in a transactional environment, including Replica-agnostic High Availability (R-HA) and Sticky High Availability (S-HA).
\item We analyze the existing literature on ACID properties (particularly, ANSI SQL Spec, Adya, Gray) to show which properties* with a focus on isolation* are achievable in HATS. Achievable properties include transactional atomicity, variants of repeatable read isolation, and session guarantees. We also describe unavailable properties, like preventing Lost Update and Write Skew. We believe that this is the first such characterization in the literature.
\item We further motivate the study of HATS via the analysis of existing concurrency control algorithms in a highly available environment as well as through deployment of an experimental database prototype on public cloud infrastructure.
\end{itemize}

### 2013-03-20 writeup/motivation.tex

For networks that do not partition, maintaining semantic guarantees
and available operation is
simple~\cite{stonebraker2010errors}. However, this is often not the
case: according to James Hamilton, Vice President and Distinguished
Engineer on the Amazon Web Service team, ``network partitions should
be rare but net gear continues to cause more issues than it
should''~\cite{hamilton-partitions}. Anecdotal evidence from companies
such as Google~\cite{dean-keynote} confirm Hamilton's statement, while
a host of recent and highly public cloud computing incidents highlight
partition behavior. In April 2011, a network misconfiguration led to a
twelve-hour long series of outages across the Amazon EC2 and RDS
services~\cite{amazon-netpartition}. Subsequent misconfigurations and
partial failures such as another EC2 failure in October 2012 have led
to full site outages for web services like Reddit, Foursquare, and
Heroku~\cite{ec2-downsites}. At a global scale, hardware
failures---like the 2011 outages in Level3's backbones in North
America and Europe due a 2011 bug in Juniper's
routers~\cite{juniper-partition}---and misconfigurations---such as BGP
faults introduced by Pakistan in 2008~\cite{pakistan-youtube} and
academics in 2010~\cite{research-experiment-partition}---can cause
widespread---even Internet-wide---partitions. Many of our discussions
with practitioners---especially those operating on public cloud
infrastructure---confirm that partitions are a reality.

>>>>

For networks that do not partition, maintaining semantic guarantees
and available operation is
simple~\cite{stonebraker2010errors}. However, this is often not the
case: according to James Hamilton, Vice President and Distinguished
Engineer on the Amazon Web Service team, ``network partitions should
be rare but net gear continues to cause more issues than it
should''~\cite{hamilton-partitions}. Anecdotal evidence from companies
such as Google~\cite{dean-keynote} confirm Hamilton's statement, while
a host of recent and highly public cloud computing incidents highlight
partition behavior. As an example of an actual system rearchitecture,
less than one year after its announcement, Yahoo!'s PNUTS developers
explicitly added a support for weaker, highly available operation,
explaining that ``strict adherence [to strong consistency] leads to
difficult situations under network partitioning or server
failures...in many circumstances, applications need a relaxed
approach''~\cite{pnuts-update}.

Provider outages are often publicly observable. In April 2011, a
network misconfiguration led to a twelve-hour long series of outages
across the Amazon EC2 and RDS
services~\cite{amazon-netpartition}. Subsequent misconfigurations and
partial failures such as another EC2 failure in October 2012 have led
to full site outages for web services like Reddit, Foursquare, and
Heroku~\cite{ec2-downsites}. At a global scale, hardware
failures---like the 2011 outages in Level3's backbones in North
America and Europe due a 2011 bug in Juniper's
routers~\cite{juniper-partition}---and misconfigurations---such as BGP
faults introduced by Pakistan in 2008~\cite{pakistan-youtube} and
academics in 2010~\cite{research-experiment-partition}---can cause
widespread---even Internet-wide---partitions. Many of our discussions
with practitioners---especially those operating on public cloud
infrastructure---confirm that partitions are a reality.
